---
title: "Biostat M280 Homework 3"
subtitle: Due Mar 1 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use tidyverse and bash to explore following two data sets.

## Q1 LA City Employee Payroll

The `/home/m280data/la_payroll/City_Employee_Payroll.csv` file on teaching server contains payroll information of LA City employees in years 2013-2018. It was downloaded from [LA City Controller's Office](https://controllerdata.lacity.org/Payroll/City-Employee-Payroll/pazn-qyym). Make a Shiny app to facilitate visualization of this data. 

1. For efficiency of the Shiny app, you should first pre-process, pare down, tidy, and save the data, e.g., as a compressed RDS file, to be used in the app.
```{bash, eval = (Sys.info()[["sysname"]] == "Linux")}
ln -sf /home/m280data/la_payroll/City_Employee_Payroll.csv
```



0. **Total payroll by LA City**. Visualize the total LA City payroll of each year, with breakdown into base pay, overtime pay, and other pay.

0. **Who earned most?** Visualize the payroll information (total payment with breakdown into base pay, overtime pay, and other pay, Department, Job Title) of the top $n$ highest paid LA City employees in a specific year. User specifies $n$ (default 10) and year (default 2017).

0. **Which departments earn most?** Visualize the mean or median payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ earning departments. User specifies $n$ (default 5), year (default 2017), and method (mean or median, default median).

0. **Which departments cost most?** Visualize the total payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ expensive departments. User specifies $n$ (default 5) and year (default 2017).

0. Visualize any other information you are interested in.

0. Publish your Shiny app to <https://www.shinyapps.io> and share the link.

## Q2 LA City Parking War

The SQLite database `/home/m280data/la_parking/LA_Parking_Citations.sqlite` on teaching server contains information about parking tickets in LA City. It was downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). Connect to the database and answer following questions using plots and summary statistics. In this exercise, you are **not** allowed to load whole data into memory. Use the _transform in database, plot in R_ strategy.

```{r}
library(DBI)
library(RSQLite)
library(tidyverse)
con = dbConnect(RSQLite::SQLite(), "/home/m280data/la_parking/LA_Parking_Citations.sqlite")

#Name of the table in database is latix#
dbListTables(con)
#Use dplyr package
latix_sql <- dplyr::tbl(con, "latix")
#Check all variables in the table#
latix_sql %>% print(width = Inf)
```

1. How many tickets are in this data set? Which time period do these tickets span? Which years have most data?
```{r}
latix_sql %>% filter(!is.na(Ticket_number)) %>%
  count()
```

```{r}
timespan <- latix_sql %>% filter(!is.na(Issue_Year), !is.na(Issue_Month),
                             !is.na(Issue_Day), !is.na(Issue_Hour), 
                             !is.na(Issue_Minute)) %>%
  arrange(desc(Issue_Year), desc(Issue_Month), desc(Issue_Day), 
          desc(Issue_Hour), desc(Issue_Minute)) %>% collect() %>%
  slice(c(1, n())) 
timespan
```

** The time span is from 2015-07-02 01:00 to 2019-01-25 23:58 **

```{r}
latix_sql %>% filter(!is.na(Issue_Year)) %>% group_by(Issue_Year) %>% 
  summarize(number = n()) %>% arrange(desc(number)) %>% head(1)
```

** 2017 has the most data **

2. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?

```{r}
#Hour#
latix_sql %>% filter(!is.na(Issue_Hour)) %>% group_by(Issue_Hour) %>% 
  summarize(number = n()) %>% arrange(desc(number)) %>% collect() %>% 
  slice(c(1, n()))
```

** You are most likely to get a ticket at hour 12, and least likely to get a ticket at hour 5 **

```{r}
#Weekdays#
latix_sql %>% filter(!is.na(Issue_Wday)) %>% group_by(Issue_Wday) %>% 
  summarize(number = n()) %>% arrange(desc(number)) %>% collect() %>% 
  slice(c(1, n()))
```

** You are most likely to get a ticket at Wednesday, and least likely to get a ticket at Monday **

```{r}
#Monthday#
latix_sql %>% filter(!is.na(Issue_Day)) %>% group_by(Issue_Day) %>% 
  summarize(number = n()) %>% arrange(desc(number)) %>% collect() %>% 
  slice(c(1, n()))
```

** You are most likely to get a ticket on 13th, and least likely to get a ticket on 31st **

```{r}
#Month#
latix_sql %>% filter(!is.na(Issue_Month)) %>% group_by(Issue_Month) %>% 
  summarize(number = n()) %>% arrange(desc(number)) %>% collect() %>% 
  slice(c(1, n()))
```

** You are most likely to get a ticket in August, and least likely to get a ticket in Feburary**

3. Which car makes received most citations?

```{r}
latix_sql %>% filter(!is.na(Make)) %>% group_by(Make) %>% 
  summarize(number = n()) %>% arrange(desc(number)) %>% collect() %>% 
  slice(1)
```

** Toyota received the most citation **


4. How many different colors of cars were ticketed? Which color attracted most tickets?

```{r}
Carcol <- latix_sql %>% filter(!is.na(Color)) %>% group_by(Color) %>%
  summarize(number = n()) %>% arrange(desc(number))
summarise(Carcol, n())
head(Carcol, 1)
```

** There are 103 different colors of cars, Black attracted the most tickets **

5. What are the most common ticket types?

```{r}
latix_sql %>% filter(!is.na(Violation_Description)) %>% 
  group_by(Violation_Description) %>% summarize(number = n()) %>% 
  arrange(desc(number)) %>% collect() %>% 
  slice(1)
```

** The most common ticket type is "No park/Street clean" **

6. How much money was collected on parking tickets in 2016, 2017 and 2018?

```{r}
latix_sql %>% filter(!is.na(Issue_Year), !is.na(Fine_amount)) %>% 
  group_by(Issue_Year) %>% summarise(Tot = sum(Fine_amount, na.rm = TRUE)) %>% 
  collect() %>% slice(c(2,3,4))
```

** Money collected on parking tickets in 2016, 2017 and 2018 are **
** $152145538, $157122489, $138875787 respectively **

0. If you've been ticketed in LA County, did you find your ticket in this data set?

0. Read the blog <http://www.brettrics.com/9-million-parking-tickets-la/> and try to reproduce plots using your data.